---
title: "milestone_5"
author: "Amy Tan"
date: "10/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreign)
library(haven)
library(dplyr)
library(tibble)
library(tidyverse)
```
<br>
---
About Page:
---
<br>

In this project, I used two datasets to figure out if there are any interesting correlations between economic variables and educational variables at the Census tract-level nationwide. How do students' broader social environments impact their academic achievement? I attempt to answer this question by getting the tract_covariates dataset from the Opportunity Insights website (https://opportunityinsights.org/data/) and the stanforddata dataset from the Stanford Education Data Archive (https://exhibits.stanford.edu/data/catalog/db586ns4974). The tract_covariates dataset describes economic and social factors of each Census tract, such as rates of intergenerational mobility and umemployment rates. The stanforddata dataset describes various academic variables for 3rd-8th graders across the nation by tract as well. It contains variables for academic achievement and student demographics. I merged these two datasets together by their FIPS code. I also descriptively analyzed each dataset to find trends within the datasets themselves, such as the correlation between race and income in the tract_covariates dataset. The graph below is one of these descriptive analyses, using the Opportunity Atlas dataset.

<br>
```{r reading_in_data, echo=FALSE}

#progress on data processing

#loading datasets in raw-data folder using terminal, and then using read_dta function for Stata files

stanforddata <- read_dta("raw-data/SEDA_county_long_CS_v21.dta")
tract_covariates <- read_dta("raw-data/tract_covariates.dta")

#getting a sense of what variables I want to include in my ggplot

view(stanforddata)
view(tract_covariates)
```

```{r merging_data, echo=FALSE}

#made the state fips code numeric instead of characters because inner_join doesn't work with characters

#learned that you have to pull the list column out of the dataframe to make it numeric

state_fips <- as.numeric(tract_covariates %>% pull(state))

#looked up add_column to create a new variable with the numeric state FIPS code

tract_edit <- add_column(tract_covariates, state_fips)

#used the same process to change the fips code in the stanforddata dataset to a numeric vector

#called it the same thing to easily inner_join later by a common variable name

state_fips <- as.numeric(stanforddata %>% pull(fips))
stan_edit <- add_column(stanforddata, state_fips)

#this merged dataset should work, but the message "error: vector memory exhausted (limit reached?)" showed up, figure out why later

#x <- inner_join(tract_edit, stan_edit, by="state_fips")

```

```{r plot, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}

#creating a filtered tibble with only the data from the state of Massachusetts that does not include NA for the two variables of interest

d <- tract_covariates %>%
  filter(!is.na(share_white2000)) %>%
  
  #filtering out NA values
  
  filter(!is.na(hhinc_mean2000)) %>%
  
  #found FIPS code for Massachusetts
  
  filter(state==25)

#remember to create a new object for ggplot

p<- ggplot(data=d, mapping= aes(x = share_white2000, y=hhinc_mean2000))+
  
#I chose geom_point because I wanted a scatterplot of all the datapoints in Massachusetts to compare how the proportions of white people might have a correlation with the mean household income
  
  geom_point(alpha=.2)+
  
#I used geom_jitter because a lot of the points were too condensed together, so this would allow me to see the individual points more clearly
  
  geom_jitter()+
  
#this creates a linear model that fits a line to the overall trend.
  
  geom_smooth()+
  
#I added some captions and titles to the ggplot to make sure I cite everything and make the graph understandable
  
  labs(x="Proportion of White People", y="Mean Household Income", title="Share of White People v. Mean Household Income", subtitle="Data from State of Massachusetts in 2000", caption="Source: Opportunity Atlas, https://opportunityinsights.org/data/")

#don't forget to print the output!
p
